---
title: "R Notebook"
output: html_notebook
---

# Q10 is aspirations 
# Q7_7_1 is javascript skills 
# Q7_7_2 is mashup skills 
# Q8_8_1 is general programming skills 

```{r}
# reading input 
df<-read.csv("input/commits_novelty.csv", header =TRUE, sep=",")
df <- df[complete.cases(df), ]  
df
```

```{r}
df$group = factor(df$group)
```


```{r}
# create new columns called log relational novelty
df$log_relational_novelty <- log(df$similarity+1) 
df$log_count <- log(df$count+1) 
df$Q7_Q7_1 <- log(df$Q7_Q7_1+1)
df$Q7_Q7_2 <- log(df$Q7_Q7_2+1)
df$Q8_Q8_1 <- log(df$Q8_Q8_1+1)
df$Q10 <- log(df$Q10+1)
df
```


```{r}
# standardizing variables for skills and aspirations. 
cols <- c("Q7_Q7_1", "Q7_Q7_2", "Q8_Q8_1", "Q10", "log_relational_novelty", "log_count")
df[cols] <- scale(df[cols])
df
```


```{r}
mod <- lm(log_count ~ factor(group), data=df)
summary(mod)
```


```{r}
mod <- lm( log_relational_novelty ~ Q10 + Q8_Q8_1 + Q7_Q7_1 + Q7_Q7_2, data = df)
summary(mod)
```

```{r}
mod <- lm( log_relational_novelty ~ log_count , data = df)
summary(mod)
```

```{r}
mod <- lm( log_relational_novelty ~ factor(group) + log_count + Q7_Q7_1 + Q7_Q7_2 + Q8_Q8_1 + Q10 , data = df)
summary(mod)
```
```{r}
mod <- lm( log_relational_novelty ~ factor(group)/stage + log_count + Q7_Q7_1 + Q7_Q7_2 + Q8_Q8_1 + Q10 , data = df)
summary(mod)
```

```{r}
mod <- lm( log_relational_novelty ~ factor(group) * stage + log_count + Q7_Q7_1 + Q7_Q7_2 + Q8_Q8_1 + Q10 , data = df)
summary(mod)
```



```{r}
# Proposed model by stepwise regression
library(stats)
mod <- lm( log_relational_novelty ~ factor(group) + log_count + Q7_Q7_2 , data = df)
summary(mod)
AIC(mod)
BIC(mod)
```

```{r}
# without the factor ( group ) and with all confounding variables 
library(stats)
mod <- lm( log_relational_novelty ~ log_count + Q7_Q7_1 + Q7_Q7_2 + Q8_Q8_1 + Q10 , data = df)
summary(mod)
AIC(mod)
BIC(mod)
```

```{r}
library(stats)
mod.1 <- lm( log_relational_novelty ~ log_count + Q7_Q7_2 , data = df)
summary(mod.1)
AIC(mod.1)
BIC(mod.1)
```

```{r}
library(stats)
mod.2 <- lm( log_relational_novelty ~ factor(group) + log_count + Q7_Q7_2 , data = df)
summary(mod.2)
AIC(mod.2)
BIC(mod.2)
```

```{r}
# model with and without groups are very different ( significant )
anova(mod.1, mod.2)
```


```{r}
library(ALSM)
step(lm(log_relational_novelty ~ factor(group) + log_count + Q7_Q7_1 + Q7_Q7_2 + Q8_Q8_1 + Q10, data=df),
method="both", trace = 1 )
```




```{r}
mod <- lm( log_relational_novelty ~ factor(group) + Q10 + Q8_Q8_1 + Q7_Q7_1 + Q7_Q7_2 , data = df)
summary(mod)
```

## Nest Phase in Group [ Linear Mixed Model ]

```{r}
# convert to nominal factor
df$user2 = factor(df$user2)
df$stage = factor(df$stage)
df$group = factor(df$group)
```

```{r}
# explore the data and their levels 
library(plyr)
ddply(df, ~ group * stage, function(data) summary(data$log_relational_novelty) )
ddply(df, ~ group * stage, summarise, log_relational_novelty.mean=mean(log_relational_novelty), log_relational_novelty.sd = sd(log_relational_novelty))
```
```{r}
# histograms for two factors
hist(df[df$group == 0 & df$stage == 1,]$log_relational_novelty)
hist(df[df$group == 0 & df$stage == 2,]$log_relational_novelty)
hist(df[df$group == 0 & df$stage == 3,]$log_relational_novelty)
hist(df[df$group == 0 & df$stage == 4,]$log_relational_novelty)
hist(df[df$group == 1 & df$stage == 1,]$log_relational_novelty)
hist(df[df$group == 1 & df$stage == 2,]$log_relational_novelty)
hist(df[df$group == 1 & df$stage == 3,]$log_relational_novelty)
hist(df[df$group == 1 & df$stage == 4,]$log_relational_novelty)
hist(df[df$group == 2 & df$stage == 1,]$log_relational_novelty)
hist(df[df$group == 2 & df$stage == 2,]$log_relational_novelty)
hist(df[df$group == 2 & df$stage == 3,]$log_relational_novelty)
hist(df[df$group == 2 & df$stage == 4,]$log_relational_novelty)
hist(df[df$group == 3 & df$stage == 1,]$log_relational_novelty)
hist(df[df$group == 3 & df$stage == 2,]$log_relational_novelty)
hist(df[df$group == 3 & df$stage == 3,]$log_relational_novelty)
hist(df[df$group == 3 & df$stage == 4,]$log_relational_novelty)
boxplot(log_relational_novelty ~ group * stage, data = df, xlab="Group.Stage", ylab="log_relational_novelty")
with(df, interaction.plot(group, stage, log_relational_novelty, ylim=c(0, max(log_relational_novelty)))) # interaction plot
```
```{r}
# library for LMM we will use on relational novelty 

library(lme4)
library(lmerTest)
library(car)
```

# set sum-to-zero contrasts for the Anova cells 

```{r}
contrasts(df$group) <= "contr.sum"
contrasts(df$stage) <= "contr.sum"
```


```{r}
# stage nested within group 
full.model = lmer( log_relational_novelty ~ group/stage + (1 | user2 ), data = df, REML = FALSE)
Anova(full.model, type=3, test.statistics="F")
full.model
```
```{r}
library(performance)

check_collinearity(full.model)


```
# variability is very much higher in individual user and than in stages/phases for relational novelty. The remaining variability of  0.203502 comes from factor other than individual users and stage. factor(group)1  has higher relational novelty than group 0 by about  0.2099 . factor(group)3 has higher relational novelty than group 0 by about 0.3765. 

```{r}
var.model = lmer( log_relational_novelty ~ factor(group) + ( 1 | user2) + ( 1 | stage), data = df)
summary(var.model)
```


```{r}
reduced.model = lm( log_relational_novelty ~ log_count + Q7_Q7_1 + Q7_Q7_2 + Q8_Q8_1 + Q10, data = df)
summary(reduced.model)
```


```{r}
full.model = lm( log_relational_novelty ~ factor(group) + log_count + Q7_Q7_1 + Q7_Q7_2 + Q8_Q8_1 + Q10, data = df)
summary(full.model)
```



```{r}
anova(reduced.model, full.model)
```


```{r}
boxplot(log_relational_novelty~ stage*group,
col=c("white","lightgray", "blue", "green"),df)
```

```{r}
check_collinearity(full.model)
```

```{r}
library(car)

vif(full.model)
```
```{r}
vif(reduced.model)
```

```{r}
library(multcomp)
library(lsmeans)
#summary(glht(full.model, lsm(pairwise ~ group / stage)), test = adjusted(type='holm'))
```

